The Mass Distribution of Stellar-Mass Black Holes

### 2010-10-20 paper.tex
As submitted to collaborators.

\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}

\newcommand{\fixme}[1]{[FIXME: #1]}
\newcommand{\Msun}{M_\odot}
\newcommand{\Mmin}{M_{\textnormal{min}}}
\newcommand{\Mmax}{M_{\textnormal{max}}}
\newcommand{\Nbin}{N_{\textnormal{bin}}}
\newcommand{\vtheta}{\vec{\theta}}

@@

\begin{abstract}
  We perform a Bayesian analysis of the mass distribution of
  stellar-mass black holes using the observed masses of 15 low-mass
  X-ray binary systems and five higher-mass, wind-fed X-ray binary
  systems.  Modeling the mass distribution both parametrically---as a
  power law, exponential, gaussian, combination of two gaussians, or
  log-normal distribution---and non-parametrically---as histograms
  with varying numbers of bins---we explore the distribution in the
  context of Markov Chain Monte Carlo calculations of model
  parameters.  We give confidence bounds on the shape of the mass
  distribution in the context of each model and compare the models
  with each other by calculating their Bayesian evidence.  The mass
  distribution of the low-mass systems is best fit by a power-law,
  while the distribution of the complete sample is best fit by the
  exponential model.  We address the existence of a ``gap'' between
  the most massive neutron stars and least massive black holes,
  finding that the best model (the power law) fitted to the low-mass
  systems only gives a minimum black-hole mass of 4.7 $\Msun$ (90\%
  confidence), while the best model (the exponential) fitted to all 20
  systems gives a minumum black-hole mass of 4.4 $\Msun$ (90\%
  confidence).  We conclude that our sample of black hole masses
  provides strong evidence of a gap between the maximum neutron star
  mass and minimum black hole mass.
\end{abstract}

@@

\begin{figure}
  \plotone{plots/masses-high}

  \caption{\label{fig:masses} The mass distributions in Figures
    \ref{fig:low-masses} and \ref{fig:high-masses} combined into a
    single distribution. Note that this combination is not a good
    estimator of the true mass distribution of black hole masses in
    the galaxy---see, e.g., \citet{Mandel2010}.  We include it here
    only to give a sense of the range and general shape of the mass
    distribution of black holes in our sample.}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \plottwo{plots/dist-parameteric}{plots/dist-non-parameteric}
  \end{center}
  \caption{\label{fig:dists} The median values of the black hole mass
    distribution, $p(M|\theta)$, at various masses implied by the
    posterior $p(\theta|d)$ for the models discussed in Sections
    \ref{subsec:parameteric-models} and
    \ref{subsec:non-parameteric-models}.  These distributions use only
    the 15 low-mass observations in Table \ref{tab:sources} (the
    complete sample is analyzed in Section \ref{sec:higher-mass}).
    Error bars span the 10\% to 90\% range.  Note that these
    ``distributions of distributions'' are not necessarily normalized,
    and need not be ``shaped'' like the underlying model
    distributions.}
\end{figure}

@@

\subsection{Priors}

@@

\subsection{Parameteric Models for the Black Hole Mass Distribution}
\label{subsec:parameteric-models}

In this subsection, we discuss the various parameteric models of the
underlying black hole mass distribution that we have analyzed using
the low-mass data from Section \ref{sec:systems}.  (Analysis of these
models including the higher-mass systems appears in Section
\ref{sec:higher-mass}.)

\subsubsection{Power-Law Models}
\label{subsubsec:power-law}

@@

The marginalized distribution for $\alpha$ is broad, with
\begin{equation}
  -11.8 < \alpha < 6.8
\end{equation}
enclosing 90\% of the probability.  We have $p(\alpha < 0) = 0.6$.
The median value is $\alpha = -3.35$.  The distribution of the minimum
mass is also quite interesting, but we defer this discussion until
Section \ref{sec:minimum-mass}, where we discuss the distribution for
the minimum black hole masses in the context of all our models.

\begin{figure}
  \begin{center}
    \plotone{plots/power-law}
  \end{center}
  \caption{\label{fig:power-law} A histogram of the MCMC samples for
    $\alpha$ from the power-law distribution defined in Equation
    \eqref{eq:power-law-dist}.  The histogram is an approximation to
    the one-dimensional marginalized PDF in Equation
    \eqref{eq:alpha-pdf}.  The favored value is near $\alpha = -4$.
    The fraction of the distribution with $\alpha < 0$ is 98.7\%.  The
    5\% quantile is $-7.66$, the median is $-4.28$, and the 95\%
    quantile is $-1.36$.}
\end{figure}

\subsubsection{Decaying Exponential}
\label{subsubsec:exponential}

Theoretical arguments in \citet{Fryer2001} suggest that the black-hole
mass distribution may be well-represented by a decaying exponential
with a minimum mass:
\begin{equation}
  \label{eq:exp-def}
  p(m|\theta) = p(m|\{\Mmin, M_0\}) = 
  \begin{cases}
    \frac{e^{\frac{\Mmin}{M_0}}}{M_0} \exp \left[ - \frac{m}{M_0}
    \right] & \Mmin \leq m \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}
We choose again a broad, flat prior
\begin{equation}
  p(\{\Mmin, M_0 \}) = 
  \begin{cases}
    \frac{1}{40^2} & 0 \leq \Mmin, M_0 \leq 40 \\
    0 & \textnormal{otherwise}
  \end{cases},
\end{equation}
with $\Mmin$ and $M_0$ measured in solar masses.

%Figure \ref{fig:exp-M0}
 displays the marginalized posterior
distribution for the scale mass of the exponential, $M_0$.  The median
scale mass is $M_0 = 3.00$, and $1.57 \leq M_0 \leq 5.71$ with 90\%
confidence.  We defer discussion of the distribution of $\Mmin$ until
Section \ref{sec:minimum-mass}.  Refer to Figure \ref{fig:dists}
for the resulting shape of the distribution of distributions.

% \begin{figure}
%   \begin{center}
%     \plotone{plots/exp-M0}
%   \end{center}
%   \caption{\label{fig:exp-M0} The distribution of scale masses, $M_0$,
%     measured in units of a solar mass for the exponential underlying
%     mass distribution defined in Equation \eqref{eq:exp-def}.  The
%     median scale mass is $M_0 = 3.00$; with 90\% confidence $1.57 \leq
%     M_0 \leq 5.71$. }
% \end{figure}

\subsubsection{Gaussian and Two-Gaussian Models}
\label{subsubsec:gaussian}

The prototypical single-peaked probability distribution is a
Gaussian:
\begin{equation}
  p(m|\theta) = p(m|\{\mu, \sigma\}) = \frac{1}{\sigma \sqrt{2\pi}}
  \exp\left[ - \left(\frac{m - \mu}{\sqrt{2} \sigma} \right)^2 \right].
\end{equation}
We choose a broad, flat prior 
\begin{equation}
  \label{eq:gaussian-def}
  p(\{\mu,\sigma\}) = 
  \begin{cases}
    \frac{1}{40^2} & 0 \leq \mu, \sigma \leq 40 \\
    0 & \textnormal{otherwise}
  \end{cases},
\end{equation}
where both $\mu$ and $\sigma$ are measured in solar masses.  Figure
\ref{fig:gaussian} shows the resulting marginalized distributions for
the parameters $\mu$ and $\sigma$.  We constrain $\mu$ in $7.02 \leq
\mu \leq 9.23$ with 90\% confidence.  (Refer to Figure \ref{fig:dists}
for the shape of the resulting distribution of distributions.)

\begin{figure}
  \begin{center}
    \plotone{plots/gaussian}
  \end{center}
  \caption{\label{fig:gaussian} Marginalized posterior distributions
    for the mean $\mu$ and standard deviation $\sigma$ (both in solar
    masses) for the Gaussian underlying mass distribution defined in
    Equation \eqref{eq:gaussian-def}.  The peak of the Gaussian,
    $\mu$, is constrained in $7.02 \leq \mu \leq 9.23$ with 90\%
    confidence.}
\end{figure}

To look for a second peak in the black-hole mass distribution, we also
examined a two-Gaussian model:
\begin{multline}
  p(m|\theta) = p(m|\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = \\
  \frac{\alpha}{\sigma_1 \sqrt{2\pi}} \exp\left[ - \left( \frac{m -
        \mu_1}{\sqrt{2}\sigma_1} \right)^2 \right] + \frac{1-\alpha}{\sigma_2 \sqrt{2\pi}} \exp\left[ - \left( \frac{m -
        \mu_2}{\sqrt{2}\sigma_2} \right)^2 \right],
\end{multline}
where we restrict $\mu_1 < \mu_2$ with our broad, flat prior:
\begin{equation}
  p(\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = 
  \begin{cases}
    2 \frac{1}{40^4} & 0 \leq \mu_1 \leq \mu_2 \leq 40,
    \quad 0 \leq \sigma_1, \sigma_2 \leq 40, \quad 0 \leq \alpha \leq
    1 \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}
The one-Gaussian model is a subset of the Gaussian model: when
$\alpha$ is small, then the presence of the first Gaussian is
irrelevant, and the second gives the same mass distribution as a
single Gaussian; when $\alpha$ is large, the reverse situation again
reproduces the single-Gaussian model; and when $\alpha$ is in the
middle of its range, the two means and standard deviations can be
nearly equal, reproducing again the single-Gaussian model.  This is
precisely what we observe from our MCMC samples in the two-Gaussian
model.  In Section \ref{sec:model-selection}, we will quantify more
precisely how much redundancy we have in this model relative to the
one-Gaussian model, but we should emphasize here that there is no
evidence of a significant contribution from a second peak in the
two-Gaussian model.

@@

\subsection{Non-Parameteric Models for the Black Hole Mass Distribution}
\label{subsec:non-parameteric-models}

The previous subsection discussed models for the underlying black hole
mass distribution that assumed particular parameterized shapes for the
distribution.  In this subsection, we will discuss models that do not
assume a priori a shape for the black hole mass distribution.  The
fundamental non-parameteric distribution in this section is a
histogram with some number of bins, $\Nbin$.  Such a distribution is
piecewise-constant in $m$.

@@

We choose to represent a histogram mass distribution with $\Nbin$ bins
by allocating a fixed total probability, $1/\Nbin$, to each bin.  The
lower and upper bounds for each bin are allowed to vary; when these
are close to each other (i.e.\ the bin is narrow), the distribution
will have a large value, and conversely when the bounds are far from
each other.  We assume that the non-zero region of the distribution is
contiguous, so we can represent the boundaries of the bins as a
non-decreasing array of masses, $w_0 \leq w_1 \leq \ldots \leq
w_{\Nbin}$, with $w_0$ the minimum and $w_{\Nbin}$ the maximum mass for
which the distribution has support.  This gives the distribution
\begin{equation}
  \label{eq:hist-def}
  p(m|\theta) = p(m|\{w_0, \ldots, w_{\Nbin}\}) = 
  \begin{cases}
    0 & m < w_0 \textnormal{ or } w_{\Nbin} \leq m \\
    \frac{1}{\Nbin} \frac{1}{w_{i+1} - w_i} & w_i \leq m < w_{i+1}
  \end{cases}.
\end{equation}

@@

The median values of the histogram mass distributions that result from
the MCMC samples of the posterior distribution for the $w_i$
parameters for one-, two-, three-, four-, and five-bin histogram
models are shown in Figure \ref{fig:dists}.  As the number of bins
increases, the models are better able to capture features of the mass
distribution.  (But note that in Section \ref{sec:model-selection}, we
will find that the one-bin histogram---which is a flat distribution
between a maximum and minimum mass---is the best model from this
group.)  We defer discussion of the distribution of the minimum mass
from the histogram distributions until Section \ref{sec:minimum-mass}.

>>>>

%\usepackage{hyperref}
\usepackage{amssymb}
\usepackage{amsmath}

\newcommand{\fixme}[1]{[FIXME: #1]}
\newcommand{\Msun}{M_\odot}
\newcommand{\Mmin}{M_{\textnormal{min}}}
\newcommand{\Mmax}{M_{\textnormal{max}}}
\newcommand{\Nbin}{N_{\textnormal{bin}}}
\newcommand{\vtheta}{\vec{\theta}}
\newcommand{\order}[1]{\mathcal{O}\left( #1 \right)}

@@

\begin{abstract}
  We perform a Bayesian analysis of the mass distribution of
  stellar-mass black holes using the observed masses of 15 low-mass
  X-ray binary systems undergoing Roche lobe overflow and five
  higher-mass, wind-fed X-ray binary systems.  Modeling the mass
  distribution both parametrically---as a power law, exponential,
  gaussian, combination of two gaussians, or log-normal
  distribution---and non-parametrically---as histograms with varying
  numbers of bins---we explore the distribution in the context of
  Markov Chain Monte Carlo calculations of model parameters.  We give
  confidence bounds on the shape of the mass distribution in the
  context of each model and compare the models with each other by
  calculating their Bayesian evidence.  The mass distribution of the
  low-mass systems is best fit by a power-law, while the distribution
  of the complete sample is best fit by the exponential and power-law
  models (but with significantly different parameters than the
  lower-mass sample).  We address the existence of a ``gap'' between
  the most massive neutron stars and least massive black holes,
  finding that the best model (the power law) fitted to the low-mass
  systems only gives a minimum black-hole mass of 4.3 $\Msun$ (90\%
  confidence), while the best models (the exponential and power-law)
  fitted to all 20 systems gives a minumum black-hole mass of 4.5 and
  5.2 $\Msun$ (90\% confidence), respectively.  We conclude that our
  sample of black hole masses provides strong evidence of a gap
  between the maximum neutron star mass and minimum black hole mass.
\end{abstract}

@@

\begin{figure}
  \plottwo{plots/masses}{plots/masses-high}

  \caption{\label{fig:masses} The mass distributions in Figure
    \ref{fig:low-masses} (left) and both \ref{fig:low-masses} and
    \ref{fig:high-masses} (right) combined into a single
    distribution. Note that this combination is not a good estimator
    of the true mass distribution of black hole masses in the
    galaxy---see, e.g., \citet{Mandel2010}.  We include it here only
    to give a sense of the range and general shape of the mass
    distribution of black holes in our sample.}
\end{figure}

@@

\begin{figure}
  \begin{center}
    \plottwo{plots/dist-parametric}{plots/dist-non-parametric}
  \end{center}
  \caption{\label{fig:dists} The median values of the black hole mass
    distribution, $p(M|\theta)$, at various masses implied by the
    posterior $p(\theta|d)$ for the models discussed in Sections
    \ref{sec:parametric-models} and
    \ref{sec:non-parametric-models}.  These distributions use only
    the 15 low-mass observations in Table \ref{tab:sources} (the
    complete sample is analyzed in Section \ref{sec:higher-mass}).
    Error bars span the 10\% to 90\% range.  Note that these
    ``distributions of distributions'' are not necessarily normalized,
    and need not be ``shaped'' like the underlying model
    distributions.}
\end{figure}

@@

\subsection{Priors}
\label{sec:priors}

@@

\subsection{Parametric Models for the Black Hole Mass Distribution}
\label{sec:parametric-models}

In this subsection, we discuss the various parametric models of the
underlying black hole mass distribution that we have analyzed using
the low-mass data from Section \ref{sec:systems}.  (Analysis of these
models including the higher-mass systems appears in Section
\ref{sec:higher-mass}.)  Table \ref{tab:low-mass-parametric} gives
quantiles of the marginalized parameter distributions of the
parametric models.

\begin{table}
  \begin{center}
    \begin{tabular}{|l|c|c|c|c|c|c|}
      \hline
      Model & Parameter & 5\% & 15\% & 50\% & 85\% & 95\% \\
      \hline \hline
      Power Law (Equation \eqref{eq:power-law-dist}) & $\Mmin$ & 
      1.2786 &  4.1831 &  6.1001 &  6.5011 &  6.6250 \\
      \hline
       & $\Mmax$ & 8.5578 &  8.9214 & 23.3274 & 36.0002 & 38.8113 \\
       \hline
       & $\alpha$ & -12.4191 & -10.1894 & -6.3861 &  2.8476 &  5.6954 \\
       \hline \hline
       Decaying Exponential (Equation \eqref{eq:exp-def}) & $\Mmin$ & 
       5.0185 &  5.4439 &  6.0313 &  6.3785 &  6.5316 \\
       \hline
       & $M_0$ & 0.7796 &  0.9971 & 1.5516 &  2.4635 &  3.2518 \\
       \hline \hline
       Gaussian (Equation \eqref{eq:gaussian-def}) & $\mu$ & 
       6.6349 &  6.9130 &  7.3475 & 7.7845 & 8.0798 \\
       \hline
       & $\sigma$ & 0.7478 &  0.9050  & 1.2500 &  1.7335 & 2.1134 \\
       \hline \hline
       Two Gaussian (Equation \eqref{eq:two-gaussian-def}) & $\mu_1$ & 
       5.4506 &  6.3877 &  7.1514 &  7.6728  & 7.9803 \\
       \hline
       & $\mu_2$ & 7.2355 &  7.7387 & 12.3986 & 25.2456 & 31.4216 \\
       \hline
       & $\sigma_1$ & 0.3758 &  0.7626 &  1.2104 &  1.7981 &  2.3065 \\
       \hline
       & $\sigma_2$ & 0.2048 & 0.6421 & 1.9182 &  5.2757  & 7.2625 \\
       \hline
       & $\alpha$ & 0.0983 &  0.3526 & 0.8871 &  0.9792 &  0.9936 \\
       \hline \hline
       Log Normal (Equation \eqref{eq:log-normal-def}) & $\langle M \rangle$ & 
       6.7619 &  7.0122 &  7.4336  &  7.9159  &  8.2942 \\
       \hline 
       & $\sigma_M$ & 0.7292  &  0.8920  & 1.2704  &  1.8695  &  2.4069 \\
       \hline
    \end{tabular}
  \end{center}
  \caption{\label{tab:low-mass-parametric} Quantiles of the
    marginalized distribution for each of the parameters in the models
    discussed in Section \ref{sec:parametric-models}.  We indicate
    the 5\%, 15\%, 50\% (median), 85\%, and 95\% quantiles.  The
    marginalized distribution can be misleading when there are strong
    correlations between variables.  For example, while the
    marginalized distributions for the power law parameters are quite
    broad, the distribution of mass distributions implied by the power
    law MCMC samples is similar to the other models.  This occurs in
    spite of the broad marginalized distributions because of the
    correlations between the slope and limits of the power law
    discussed in Section \ref{sec:power-law}.}
\end{table}

\subsubsection{Power-Law Models}
\label{sec:power-law}

@@

\begin{figure}
  \begin{center}
    \plotone{plots/power-law}
  \end{center}
  \caption{\label{fig:power-law} A histogram of the marginalized
    distribution for the three parameters $\Mmin$, $\Mmax$, and
    $\alpha$ from the power-law model.  The marginalized distribution
    for $\alpha$ is broad, with $-11.8 < \alpha < 6.8$ enclosing 90\%
    of the probability.  We have $p(\alpha < 0) = 0.6$; the median
    value is $\alpha = -3.35$.  The broad distribution for $\alpha$
    (and the other parameters) is due to correlations between the
    parameters discussed in the main text; see Figure
    \ref{fig:power-law-2D}.}
\end{figure}

The marginalized distribution for $\alpha$ is broad, with
\begin{equation}
  -11.8 < \alpha < 6.8
\end{equation}
enclosing 90\% of the probability (excluding 5\% on each side).  We
have $p(\alpha < 0) = 0.6$.  The median value is $\alpha = -3.35$.
The broadness of the marginalized distribution for $\alpha$ comes from
the need to match the relatively narrow peak in the mass distribution
apparent in Figure \ref{fig:masses}.  When $\alpha$ is negative, the
resulting mass distribution slopes down; $\Mmin$ is constrained to be
near the lowest mass of the observed black holes, while $\Mmax$ is
essentially irrelevant.  Conversely, when $\alpha$ is positive and the
mass distribution slopes up, $\Mmax$ must be close to the largest mass
observed, while $\Mmin$ is essentially irrelevant.  Figure
\ref{fig:power-law-2D} illustrates this effect, showing the
correlations between $\alpha$ and $\Mmin$ and $\alpha$ and $\Mmax$.
When we include the higher-mass systems in the analysis, the long tail
will eliminate this effect by bringing both $\Mmin$ and $\Mmax$ into
play for all values of $\alpha$.

\begin{figure}
  \begin{center}
    \plotone{plots/power-law-2D}
  \end{center}
  \caption{\label{fig:power-law-2D} MCMC samples in the $\Mmin,
    \alpha$ (top) and $\Mmax, \alpha$ (bottom) planes for the
    power-law model discussed in Section \ref{sec:power-law}.  The
    correlations between $\alpha$ and the power law bounds discussed
    in the text are apparent: when $\alpha$ is positive, the mass
    distribution slopes upward and $\Mmax$ is constrained to be near
    the maximum observed mass while $\Mmin$ is unconstrained.  When
    $\alpha$ is negative, the mass distribution slopes down and
    $\Mmin$ is constrained to be near the lowest mass observed, while
    $\Mmax$ is unconstrained. }
\end{figure}

\subsubsection{Decaying Exponential}
\label{sec:exponential}

\citet{Fryer2001} suggest that the black-hole mass distribution may be
well-represented by a decaying exponential with a minimum mass:
\begin{equation}
  \label{eq:exp-def}
  p(M|\vtheta) = p(M|\{\Mmin, M_0\}) = 
  \begin{cases}
    \frac{e^{\frac{\Mmin}{M_0}}}{M_0} \exp \left[ - \frac{M}{M_0}
    \right] & \Mmin \leq M \\
    0 & \textnormal{otherwise}
  \end{cases}.
\end{equation}
We choose a prior for this model where $\Mmin$ is uniform between 0
and $40 \Msun$.  For each $\Mmin$, we choose $M_0$ uniformly within a
range ensuring that $40\Msun$ is at least two scale masses above the
cutoff: $40\Msun \geq \Mmin + 2 M_0$.  This ensures that the majority
of the mass probability lies in the range $0 \leq M \leq 40\Msun$.
The resulting prior is
\begin{equation}
  p(\vtheta) = p(\{\Mmin, M_0\}) = 
  \begin{cases}
    \frac{4}{40^2} & 0 \leq \Mmin \leq 40, \quad 0 < M_0, \quad \Mmin
    + 2 M_0 \leq 40, \\
    0 & \textnormal{otherwise}
  \end{cases}
\end{equation}

Figure \ref{fig:exp-marginal} displays the marginalized posterior
distribution for the scale mass of the exponential, $M_0$, and the
cutoff mass, $\Mmin$.  The median scale mass is $M_0 = 1.55$, and
$0.78 \leq M_0 \leq 3.25$ with 90\% confidence.  Figure
\ref{fig:exp-2D} displays the MCMC samples in the $\Mmin$, $M_0$ plane
for this model.  There is a small correlation between smaller $\Mmin$
and larger $M_0$, which is driven by the need to widen the
distribution to encompass the peak of the mass measurements in Figure
\ref{fig:low-masses} when the minimum mass is smaller.

\begin{figure}
  \begin{center}
    \plotone{plots/exp-cutoff}
  \end{center}
  \caption{\label{fig:exp-marginal} The distribution of scale masses,
    $M_0$, and minimum masses, $\Mmin$, both measured in units of a
    solar mass for the exponential underlying mass distribution
    defined in Equation \eqref{eq:exp-def}.  The median scale mass is
    $M_0 = 1.55$, and $0.78 \leq M_0 \leq 3.25$ with 90\%
    confidence.}
\end{figure}

\begin{figure}
  \begin{center}
    \plotone{plots/exp-cutoff-2d}
  \end{center}
  \caption{\label{fig:exp-2D} The MCMC samples in the $\Mmin$, $M_0$
    plane for the decaying exponential underlying mass distribution
    model.  The slight correlation between smaller $\Mmin$ and larger
    $M_0$ is driven by the need to widen the mass distribution to
    encompass the peak of the measurements in Figure
    \ref{fig:low-masses} when the minimum mass decreases.}
\end{figure}

\subsubsection{Gaussian and Two-Gaussian Models}
\label{sec:gaussian}

The mass distribution in Figure \ref{fig:low-masses} appears to have a
single peak.  The prototypical single-peaked probability distribution
is a Gaussian:
\begin{equation}
  \label{eq:gaussian-def}
  p(M|\vtheta) = p(M|\{\mu, \sigma\}) = \frac{1}{\sigma \sqrt{2\pi}}
  \exp\left[ - \left(\frac{M - \mu}{\sqrt{2} \sigma} \right)^2 \right].
\end{equation}
We use a prior on the mean mass, $\mu$, and the standard deviation,
$\sigma$, that ensures that the majority of the mass distribution lies
below $40 \Msun$:
\begin{equation}
  \label{eq:gaussian-prior-def}
  p(\{\mu,\sigma\}) = 
  \begin{cases}
    \frac{8}{40^2} & 0 \leq \mu \leq 40, \quad \sigma \geq 0, \quad
    \mu + 2\sigma \leq 40 \\
    0 & \textnormal{otherwise}
  \end{cases},
\end{equation}
where both $\mu$ and $\sigma$ are measured in solar masses.  Figure
\ref{fig:gaussian} shows the resulting marginalized distributions for
the parameters $\mu$ and $\sigma$.  We constrain the peak of the
Gaussian between $6.63 \leq \mu \leq 8.08$ with 90\% confidence.
(Refer to Figure \ref{fig:dists} for the shape of the resulting
distribution of distributions.)

\begin{figure}
  \begin{center}
    \plotone{plots/gaussian}
  \end{center}
  \caption{\label{fig:gaussian} Marginalized posterior distributions
    for the mean $\mu$ and standard deviation $\sigma$ (both in solar
    masses) for the Gaussian underlying mass distribution defined in
    Equation \eqref{eq:gaussian-def}.  The peak of the Gaussian,
    $\mu$, is constrained in $6.63 \leq \mu \leq 8.08$ with 90\%
    confidence.}
\end{figure}

Though we do not expect to find a second peak in the low-mass
distribution, we may find evidence of one when exploring the combined
low- and higher-mass samples.  To look for a second peak in the
black-hole mass distribution, we use a two-Gaussian model:
\begin{multline}
  \label{eq:two-gaussian-def}
  p(M|\vtheta) = p(M|\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = \\
  \frac{\alpha}{\sigma_1 \sqrt{2\pi}} \exp\left[ - \left( \frac{M -
        \mu_1}{\sqrt{2}\sigma_1} \right)^2 \right] + \frac{1-\alpha}{\sigma_2 \sqrt{2\pi}} \exp\left[ - \left( \frac{M -
        \mu_2}{\sqrt{2}\sigma_2} \right)^2 \right].
\end{multline}
The probability is a linear combination of two Gaussians with weight
$\alpha$.  We restrict $\mu_1 < \mu_2$ and also impose combined
conditions on $\mu_i$ and $\sigma_i$ that ensure that most of the mass
probability lies below $40 \Msun$ with the prior 
\begin{equation}
  p(\{\mu_1, \mu_2, \sigma_1, \sigma_2, \alpha\}) = 
  \begin{cases}
    2 p(\{\mu_1, \sigma_1\}) p(\{\mu_2, \sigma_2\}) & \mu_1 \leq
    \mu_2, \quad 0 \leq \alpha \leq 1 \\
    0 & \textnormal{otherwise}
  \end{cases},
\end{equation}
where the single-Gaussian prior, $p(\{\mu_i, \sigma_i\})$, is defined
in Equation \eqref{eq:gaussian-prior-def}.

Figure \ref{fig:two-gaussian} shows the marginalized distributions for
the two-Gaussian model parameters from our MCMC runs.  We find $\alpha
> 0.8$ with 62\% probability, clearly favoring the first Gaussian.
The distribution for $\mu_1$ and $\sigma_1$ are similar to those of
the single Gaussian displayed in Figure \ref{fig:gaussian}, indicating
that the first Gaussian has found the peak mass samples in Figure
\ref{fig:masses}.  The second Gaussian's parameter distributions are
much broader.  The second Gaussian appears to be sampling the tail of
the mass samples (see Figure \ref{fig:masses}).  In spite of the
ability to match a more complicated distribution, we find that this
model is strongly disfavored relative to the single-Gaussian model for
this dataset: $p(\textnormal{Gaussian}|d) / p(\textnormal{Two
  Gaussian}|d) \simeq 4.7$ (see Section \ref{sec:model-selection} for
discussion).

\begin{figure}
  \begin{center}
    \plotone{plots/two-gaussian}
  \end{center}
  \caption{\label{fig:two-gaussian} The marginal distributions for the
    five parameters of the two-Gaussian model.  We have $\alpha > 0.8$
    with 62\% probability, favoring the first of the two Gaussians.
    The distributions for $\mu_1$ and $\sigma_1$ are similar to those
    of the single Gaussian model displayed in Figure
    \ref{fig:gaussian}; the second Gaussian's parameter distributions
    are much broader (recall that we constrain $\mu_2 > \mu_1$).  The
    second Gaussian is attempting to fit the tail of the mass samples
    (see Figure \ref{fig:masses}).  The extra degrees of freedom in
    the distribution from the second Gaussian do not provide enough
    extra fitting power to compensate for the increase in parmeter
    space, however: the two-Gaussian model is disfavored relative to
    the single Gaussian by a factor of $4.7$ on this dataset (see
    Section \ref{sec:model-selection} for discussion).}
\end{figure}

@@

Many of the mass distributions for the systems in Figure
\ref{fig:low-masses} rise rapidly to a peak and then fall off more
slowly in a longer tail toward high masses.  So far, none of the
parameterized distributions we have discussed have this property.  In
this section, we consider a log-normal model for the underlying mass
distribution; the log-normal distribution has a rise to a peak with a
slower falloff in a long tail.

The log-normal distribution gives $\log M$ a Gaussian distribution
with mean $\mu$ and standard deviation $\sigma$:
\begin{equation}
  \label{eq:log-normal-def}
  p(M|\vtheta) = p(M|\{\mu, \sigma \}) = \frac{1}{
    \sqrt{2\pi} M \sigma} \exp\left[ -\frac{\left(\log M - \mu\right)^2}{2 \sigma^2} \right].
\end{equation}
The parameters $\mu$ and $\sigma$ are dimensionless; the mean mass
$\langle M \rangle$ and mass standard deviation $\sigma_M$ are related
to $\mu$ and $\sigma$ by
\begin{eqnarray}
  \label{eq:avg-M}
  \langle M \rangle & = & \exp\left( \mu + \frac{1}{2} \sigma^2
  \right) \\
  \label{eq:sigma-M}
  \sigma_M & = & \langle M \rangle \sqrt{\exp\left( \sigma^2 \right) - 1}.
\end{eqnarray}
For a fair comparison with the other models, we impose a prior that is
flat in $\langle M \rangle$ and $\sigma_M$.  To ensure that most of
the probability in this model occurs for masses below $40 \Msun$, we
require $\langle M \rangle + 2 \sigma_M \leq 40$, resulting in a
prior
\begin{equation}
  p(\vtheta) = p(\{\mu,\sigma\}) = 
  \begin{cases}
    \frac{4}{40^2} \left| \frac{\partial \left(\langle M \rangle,
          \sigma_M \right)}{\partial \left( \mu, \sigma \right)}
    \right| & \sigma > 0, \quad \langle M \rangle + 2 \sigma_M \leq 40
    \\
    0 & \textnormal{otherwise}
  \end{cases},
\end{equation}
where 
\begin{equation}
  \left| \frac{\partial \left(\langle M \rangle,
          \sigma_M \right)}{\partial \left( \mu, \sigma \right)}
    \right| = \frac{\exp\left( 2 \left( \mu + \sigma^2 \right)\right)
      \sigma}{\sqrt{\exp\left( \sigma^2 \right) - 1}}
\end{equation}
is the determinant of the Jacobian of the map in Equations
\eqref{eq:avg-M} and \eqref{eq:sigma-M}.

The marginal distributions for $\langle M \rangle$ and $\sigma_M$
appear in Figure \ref{fig:log-normal}.  The distributions are similar
to those for $\mu$ and $\sigma$ from the Gaussian model in Section
\ref{sec:gaussian}.

\begin{figure}
  \begin{center}
    \plotone{plots/log-normal}
  \end{center}
  \caption{\label{fig:log-normal} Marginalized distributions of the
    mean mass, $\langle M \rangle$, and standard deviation of the
    mass, $\sigma_M$, for the log-normal model in Section
    \ref{sec:log-normal}.  The distributions are similar to the
    distributions of $\mu$ and $\sigma$ in the Gaussian model of
    Section \ref{sec:gaussian}.}
\end{figure}

\subsection{Non-Parametric Models for the Black Hole Mass Distribution}
\label{sec:non-parametric-models}

The previous subsection discussed models for the underlying black hole
mass distribution that assumed particular parameterized shapes for the
distribution.  In this subsection, we will discuss models that do not
assume a priori a shape for the black hole mass distribution.  The
fundamental non-parametric distribution in this section is a
histogram with some number of bins, $\Nbin$.  Such a distribution is
piecewise-constant in $M$.

@@

We choose to represent a histogram mass distribution with $\Nbin$ bins
by allocating a fixed probability, $1/\Nbin$, to each bin.  The lower
and upper bounds for each bin are allowed to vary; when these are
close to each other (i.e.\ the bin is narrow), the distribution will
have a large value, and conversely when the bounds are far from each
other.  We assume that the non-zero region of the distribution is
contiguous, so we can represent the boundaries of the bins as a
non-decreasing array of masses, $w_0 \leq w_1 \leq \ldots \leq
w_{\Nbin}$, with $w_0$ the minimum and $w_{\Nbin}$ the maximum mass
for which the distribution has support.  This gives the distribution
\begin{equation}
  \label{eq:hist-def}
  p(M|\theta) = p(M|\{w_0, \ldots, w_{\Nbin}\}) = 
  \begin{cases}
    0 & M < w_0 \textnormal{ or } w_{\Nbin} \leq M \\
    \frac{1}{\Nbin} \frac{1}{w_{i+1} - w_i} & w_i \leq M < w_{i+1}
  \end{cases}.
\end{equation}

@@

The median values of the histogram mass distributions that result from
the MCMC samples of the posterior distribution for the $w_i$
parameters for one-, two-, three-, four-, and five-bin histogram
models are shown in Figure \ref{fig:dists}.  Table
\ref{tab:low-mass-non-parametric} gives quantiles of the marginalized
bin boundary distributions for the histogram models.
